---
title: "lab 6"
author: "Mari Herbst"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank based test example

Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t-test)

H0: Samples are drawn from populations with equal ranks (or medians)
HA: Samples are drawn from populations where ranks (or medians) are not equal

### Create samples

'set.seed()' to create "pseudorandom, but reproducible samples"

```{r}
set.seed(1414) # so we all have the same set in the whole class
gp_1 <- sample.int(20, size = 15, replace = TRUE) # random sample of integers from 1-20, 15 items, #s repeatable

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a quick look at the distributions using the 'hist()' function

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so we can compare the results:

```{r}
ttest <- t.test(gp_1, gp_2) # unpaired 2-sample t-test

# if these data were paired, I can just add the argument 'paired = TRUE' within the t.test() function
```

This p-value means that there is a `r round(ttest$p.value, 2)` probability of finding samples with means at least this different if they were drawn from populations with the same mean

[A more comprehensive explanation of differences between groups...actual means, data spread, effect size]...there is no significant different (t(df) = t-statistic value, p = ).

### Mann-Whitney U rank based test

We use the 'wilcox.test()' function to perform a 2-sample rank based test.

```{r}
mwu <- wilcox.test(gp_1, gp_2) #unpaired rank based test
```

This p-value means that there is `r round(mwu$p.value, 2) * 100`% probability of finding samples with median at least this different if they were drawn from populations with the same median.

No significant difference in medians (Mann-Whitney U; U = `r mwu$statistic`, p = `r round(mwu$p.value, 2)`)

## Simple linear regression

Two *continuous quantitative* variables that you are trying to understand the relationship between, and you think that a linear relationship makes sense

Use the 'penguins' data from {palmerpenguins}, exploring flipper length (x) versus body mass (y)

```{r}
# use view(penguins) to look at the data
# make an exploratory scatter plot of penguin flipper length (x-axis) vs body mass (y-axis)

ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm")

# equation of regression line: y = (B_1)x + (B_0) + E

# body mass (g) = (B_1)(flipper length (mm)) + (B_0)
```

### Find the model estimates (coefficients)

Use the 'lm()' function for a linear model

body_mass_g ~ (~ = as a function of)

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

# body mass (g) = 49.69(flipper length (mm)) - 5780.08
# units: [g] = [g/mm] - [g] so 49.69  must be in [g/mm]
```

Model coefficients: `r penguin_lm$coefficients`

use indexing to pull a single coefficient value
- flipper length coefficient: `r penguin_lm$coefficients[2]`

### Use 'broom::tidy()' to get a tidy output from lm()

```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy # not finalized
```

### Use 'broom::glane()' to get output of model fit & statistical significance

```{r}
penguin_lm_glance <- glance(penguin_lm)

penguin_lm_glance
```

### Use 'equatiomatic::extract_eq()' to return model in LaTex

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)
```

### Use 'plot()' to automatically produce diagnostic plot

```{r}
plot(penguin_lm)
```

